\large How resource abundance and stochasticity affect organisms' range sizes

### 2024-06-07 writing/manuscript.Rmd
fixed ylabs of fig 5, ref to 2nd DAG, formula typo

Additional information is provided in Appendix B, including the directed acyclical graph [see fig. B6 and @mcelreath_statistical_2016] we used to infer causal the mechanisms of changes in $H$ and estimate the direct effects of $\e(R)$ and $\var(R)$ on $H$ (contrast this with fig. C5 and the empirical case study below).

@@

Separately, resource abundance and stochasticity have simple but opposing effects on organisms' range sizes: $H$ decreases with $\e(R)$ and increases with $\var(R)$. Together, the degree to which $\e(R)$ affects $H$ depends on$ $\var(R)$, and vice-versa, so organisms' responses to resource dynamics can be complex. The simulated and empirical results suggest qualitatively similar marginal effects of $\e(R)$ and $\var(R)$, but there are differences in the estimated interactive effects. In the simulated data, $\var(R)$ has little effect when $\e(R)$ is low and a strong effect when $\e(R)$ is high, while the opposite is true for the empirical data. This is due to two reasons. Firstly, the shape and symmetry of bounded distributions such as Gamma ($R > 0$) and Beta ($0 < R < 1$) distributions depend on both $\e(R)$ and $\var(R)$ (figs. A3, A4), but $\var(R)$ does not affect the shape of a Gamma distribution as much if $\e(R)$ is low (fig. B3). Secondly, and perhaps more interestingly, the simulation approach does not account for real-world adaptations to $\e(R)$ and $\var(R)$ such as selective space use, which we account for in the empirical approach. Below we discuss the strengths and limitations of each approach.

>>>>

Additional information is provided in Appendix B, including the directed acyclical graph [see fig. B6 and @mcelreath_statistical_2016] we used to infer causal the mechanisms of changes in $H$ and estimate the direct effects of $\e(R)$ and $\var(R)$ on $H$ (contrast this with fig. C3 and the empirical case study below).

@@

Separately, resource abundance and stochasticity have simple but opposing effects on organisms' range sizes: $H$ decreases with $\e(R)$ and increases with $\var(R)$. Together, the degree to which $\e(R)$ affects $H$ depends on $\var(R)$, and vice-versa, so organisms' responses to resource dynamics can be complex. The simulated and empirical results suggest qualitatively similar marginal effects of $\e(R)$ and $\var(R)$, but there are differences in the estimated interactive effects. In the simulated data, $\var(R)$ has little effect when $\e(R)$ is low and a strong effect when $\e(R)$ is high, while the opposite is true for the empirical data. This is due to two reasons. Firstly, the shape and symmetry of bounded distributions such as Gamma ($R > 0$) and Beta ($0 < R < 1$) distributions depend on both $\e(R)$ and $\var(R)$ (figs. A3, A4), but $\var(R)$ does not affect the shape of a Gamma distribution as much if $\e(R)$ is low (fig. B3). Secondly, and perhaps more interestingly, the simulation approach does not account for real-world adaptations to $\e(R)$ and $\var(R)$ such as selective space use, which we account for in the empirical approach. Below we discuss the strengths and limitations of each approach.
